---
title: "Introduction to SimEngine"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to SimEngine}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette is adapted from the <a href="https://avi-kenny.github.io/SimEngine/" target="_blank">homepage</a> of the SimEngine website. 

```{r setup}
library(SimEngine)
```

## Overview

**SimEngine** is an open-source R package for structuring, maintaining, running, and debugging statistical simulations on both local and cluster-based computing environments. Emphasis is placed on thorough documentation, and scalability.

## Getting started

The goal of many statistical simulations is to test how a new statistical method performs against existing methods. Most statistical simulations include three basic phases: (1) generate some data, (2) run one or more methods using the generated data, and (3) compare the performance of the methods. 

To briefly illustrate how these phases are implemented using **SimEngine**, we will use the example of estimating the average treatment effect of a drug in the context of a randomized controlled trial (RCT).

### 1) Load the package and create a "simulation object"

The simulation object (an R object of class *sim_obj*) will contain all data, functions, and results related to your simulation. Note that we make extensive use of the pipe operators (`%>%` and `%<>%`) from the **magrittr** package; if you have never used pipes, check out the <a href="https://magrittr.tidyverse.org" target="_blank">magrittr documentation</a>.

```{r}
library(SimEngine)
sim <- new_sim()
```

### 2) Write a function to generate some data

In **SimEngine**, functions that generate data are called *creators*. Our creator will simulate data from an RCT in which we compare a continuous outcome (e.g. blood pressure) between two groups (the "treatment group" versus the "control group"). We generate the data by looping through individuals, assigning them randomly to either the treatment group or the control group, and generating their outcome according to a simple model (note: although we use a for-loop for illustrative purposes, vectorized methods are often faster).

```{r}
create_rct_data <- function (num_patients) {
  df <- data.frame(
    "patient_id" = integer(),
    "group" = character(),
    "outcome" = double(),
    stringsAsFactors = FALSE
  )
  for (i in 1:num_patients) {
    group <- ifelse(sample(c(0,1), size=1)==1, "treatment", "control")
    treatment_effect <- ifelse(group=="treatment", -7, 0)
    outcome <- rnorm(n=1, mean=130, sd=5) + treatment_effect
    df[i,] <- list(i, group, outcome)
  }
  return (df)
}

# Test our creator function
create_rct_data(5)
```

Once you have declared your creator function, add it to your simulation object using the `add_creator()` function.

```{r}
sim %<>% add_creator(create_rct_data)
```

### 3) Code your methods

In this example, we test two different estimators of the average treatment effect. The first estimator uses the known probability of being assigned to the treatment group (0.5), whereas the second estimator uses an estimate of this probability based on the observed data. Don't worry too much about the mathematical details; the important thing is that both methods attempt to take in the dataset generated by the `create_rct_data()` function and return an estimate of the true treatment effect, which in this case is *-7*.

```{r}
estimator_1 <- function(df) {
  n <- nrow(df)
  true_prob <- 0.5
  sum_t <- sum(df$outcome * (df$group=="treatment"))
  sum_c <- sum(df$outcome * (df$group=="control"))
  return ( sum_t/(n*true_prob) - sum_c/(n*(1-true_prob)) )
}
estimator_2 <- function(df) {
  n <- nrow(df)
  est_prob <- sum(df$group=="treatment") / n
  sum_t <- sum(df$outcome * (df$group=="treatment"))
  sum_c <- sum(df$outcome * (df$group=="control"))
  return ( sum_t/(n*est_prob) - sum_c/(n*(1-est_prob)) )
}

# Test our estimator functions
df <- create_rct_data(10000)
estimator_1(df)
estimator_2(df)
```

Next, add the methods to your simulation object using the `add_method()` function.

```{r}
sim %<>% add_method(estimator_1)
sim %<>% add_method(estimator_2)
```

### 4) Set the simulation levels

Often, we want to run the same simulation multiple times (with each run referred to as a "simulation replicate"), but with certain things changed. In this example, perhaps we want to vary the number of patients and the method used to estimate the average treatment effect. We refer to the things that vary as "simulation levels". By default, **SimEngine** will run our simulation 10 times for each level combination. Below, since there are two methods and three values of num_patients, we have six level combinations and so **SimEngine** will run a total of 60 simulation replicates.

```{r}
sim %<>% set_levels(
  estimator = c("estimator_1", "estimator_2"),
  num_patients = c(50, 200, 1000)
)
```

<!-- # !!!!! add_constants()  You can also reference <a href="" target="_blank">simulation constants</a> using the variable *C*. ????? -->
<!-- A simulation constant is something that does not change across your entire set of simulations. You could just as easily hard-code values in your code, but it's often easier to have these declared at a separate section so that they can be easily modified if needed. -->

### 5) Create a simulation script

The simulation script is a function that runs a single simulation replicate and returns the results. Within a script, you can reference the current simulation level values using the variable *L*. For example, when the first simulation replicate is running, `L$estimator` will equal "estimator_1" and `L$num_patients` will equal 50. In the last simulation replicate, `L$estimator` will equal "estimator_2" and `L$num_patients` will equal 1,000. This can be used in conjunction with `use_method()` to dynamically run different methods within different simulation replicates (as is illustrated below). Your script will have access to any creators and methods that have been added to your simulation object.

```{r}
sim %<>% set_script(function() {
  df <- create_rct_data(L$num_patients)
  estimate <- use_method(L$estimator, list(df))
  return (list("estimate"=estimate))
})
```

Your script should always return a named list, although your list can be complex and contain dataframes, multiple levels of nesting, etc.

### 6) Set the simulation configuration

This controls options related to your entire simulation, such as the number of simulation replicates to run for each level combination and how to <a href="https://avi-kenny.github.io/SimEngine/parallelization/" target="_blank">parallelize</a> your code. This is discussed in detail on the <a href="https://avi-kenny.github.io/SimEngine/function-reference/set_config/" target="_blank">`set_config`</a> page. We set `num_sim` to 10, and so **SimEngine** will run a total of 60 simulation replicates (10 for each level combination).

```{r}
sim %<>% set_config(
  num_sim = 10,
  n_cores = 2,
  parallel = "outer"
)
```

### 7) Run the simulation

All 600 replicates are run at once and results are stored in the simulation object.

```{r}
sim %<>% run()
```

### 8) View and summarize results

Once the simulations have finished, use the `summarize()` function to calculate common summary statistics, such as bias, variance, MSE, and coverage.

```{r}
sim %>% summarize(
  bias = list(name="bias_ate", truth=-7, estimate="estimate"),
  mse = list(name="mse_ate", truth=-7, estimate="estimate")
)
```

In this example, we see that the MSE of estimator 1 is much higher than that of estimator 2 and that MSE decreases with increasing sample size for both estimators, as expected.

You can also directly access the results for individual simulation replicates.

```{r}
head(sim$results)
```
